---
title: "Attitudes towards outreach in particle physics"
author: "Achintya Rao"
date: "26 April 2016"
output:
    revealjs::revealjs_presentation:
        theme: serif
        highlight: pygments
        incremental: true
        center: true
        transition: slide
        reveal_options:
            slideNumber: true
        fig_caption: true
        showNotes: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
dat <- read.csv(file = "../quant_analysis/final_dataset.csv", header = TRUE)
library(dplyr)
library(plotly)
likert_units <- seq(0.0000, 5.0000, by=1)
```

---

## A bit about me

> - **Science Communicator** (2010 &mdash; present)<br />Compact Muon Solenoid (CMS) Collaboration at CERN
> - **Doctoral Student** (2014 &mdash; present)<br />UWE Science Communication Unit, Bristol

> - **MA in Science Journalism** (2009 &mdash; 2010)<br />City University, London
> - **BSc in Physics** (2005 &mdash; 2008)<br />St. Xavier’s College, Mumbai (Bombay)

<br />
<br />
<br />
<small>Twitter: [\@RaoOfPhysics](https://twitter.com/RaoOfPhysics) | GitHub: [\@RaoOfPhysics](https://github.com/RaoOfPhysics)</small>

---

<section data-background="https://cds.cern.ch/record/1363014/files/IMG_9286.JPG?subformat=icon-1440">

##

The [CMS Collaboration](http://cms.web.cern.ch) at CERN, one of the two teams that discovered the Higgs boson in 2012

- Particle physicists
- Engineers
- Students
</section>

---

<section><h1>Everything beyond this point is junk!</h1>

## Summary of cars?
```{r cars}
summary(cars)
```
</section>

---

##

```{r echo=FALSE}
set.seed(100)
d <- diamonds[sample(nrow(diamonds), 1000), ]
p <- ggplot(data = d, aes(x = carat, y = price)) +
  geom_point(aes(text = paste("Clarity:", clarity)), size = 1) +
  geom_smooth(aes(colour = cut, fill = cut)) + facet_wrap(~ cut)

(gg <- ggplotly(p))
```