---
title: "Attitudes towards outreach in particle physics"
author: "Achintya Rao"
date: "26 April 2016"
output:
    revealjs::revealjs_presentation:
        theme: serif
        highlight: pygments
        incremental: true
        center: true
        transition: slide
        reveal_options:
            slideNumber: true
        fig_caption: true
        showNotes: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
dat <- read.csv(file = "../quant_analysis/final_dataset.csv", header = TRUE)
library(dplyr)
library(plotly)
library(gridExtra)
likert_units <- seq(0.0000, 5.0000, by=1)
```

---

## A bit about me

> - **Science Communicator** (2010 &mdash; present)<br />Compact Muon Solenoid (CMS) Collaboration at CERN
> - **Doctoral Student** (2014 &mdash; present)<br />UWE Science Communication Unit, Bristol
>     - **MA in Science Journalism** (2009 &mdash; 2010)<br />City University, London
>     - **BSc in Physics** (2005 &mdash; 2008)<br />St. Xavier’s College, Mumbai (Bombay)

<br />
<br />
<small>Twitter: [\@RaoOfPhysics](https://twitter.com/RaoOfPhysics) | GitHub: [\@RaoOfPhysics](https://github.com/RaoOfPhysics)</small>

---

<section data-background="https://cds.cern.ch/record/1363014/files/IMG_9286.JPG?subformat=icon-1440">

##

- [CERN](http://home.cern), the European Laboratory for Particle Physics
    - Geneva, Switzerland
    - host laboratory for international collaborations
    - ~half of all 20,000 particle physicists

##

- The [Compact Muon Solenoid (CMS) Collaboration](http://cms.web.cern.ch)
    - one of the two teams that [discovered the Higgs boson in 2012](http://cms.web.cern.ch/news/observation-new-particle-mass-125-gev)
    - breakdown of collaborators as of June 2015
        - over 4000 physicists, engineers, students  (not including technicians and admin)
        - 183 institutes, universities, laboratories from 42 countries
        - 80+ nationalities

</section>

---

<section><h1>Data collection for quantitative analysis</h1>

## Electronic survey distributed by e-mail

- Sent to the whole collaboration via mailing list
- Three reminders sent to whole list
- Engineers reminded separately as well
- Representatives from low-response countries reminded separately

## Responses

- Over 400, but a few duplicates
- Final count: **391** valid responses
- ~10% of the collaboration

## CMS member countries

```{r}
ggplot(dat, aes(x = inst_country), y = count) +
    geom_bar()
```

## Gender

```{r}
ggplot(dat, aes(x = gender), y = count) +
    geom_bar()
```

## Age

```{r}
dat.reduced <- filter(dat, age >= "25", age < "85")

# Code below by Rik Smith-Unna (@Blahah on GitHub)

# create a vector of breaks, starting at N, then one less than each multiple
# of 10 up to the next highest multiple of ten after the maximum value
# in the data
breaks <- c(25, # start at 25, here. the c() funcion creates a vector,
            # and flattens any vectors passed in.
            # e.g try c(c(1,2,3), c(4,5,6))
            # the second argument to c() here is a sequence
            # it starts at 35
            seq(35, 
              # and ends at the maximum value rounded up to the nearest multiple
              # of ten. We get that multiple by using a feature of the
              # round() function. The first argument to round() is the
              # number to be rounded. The second is the number of digits
              # to round it to - unless the number is negative, in which
              # case it is treated as a (positive) power of ten to which
              # the number will be rounded
                round(max(dat.reduced$age + 5), -1),
              # the sequence should have step sizes of ten
                by = 10)
            # we take the sequence from 10..rounded(max) and take 1 away from
            # it. This is a vector operation - it takes 1 from every value.
            # this lets us go from 10, 20, 30... to 9, 19, 29...
            - 1)

ggplot(dat.reduced, aes(x = age)) +
  geom_histogram(breaks = breaks) +
  stat_bin(breaks = breaks,
           aes(y=..count.., label=..count..),
           geom="text",
           vjust=-0.5)
```

</section>

---

<section><h1>Attitudes</h1>

## Colleagues:

```{r}
# Importance

bar_labels <- barplot(table(dat$imp_colleagues), ylim = c(0, max(table(dat$imp_colleagues)+30)), main = "Colleagues: Importance", xlab = paste("Mean ± stddev:", format(round(mean(dat$imp_colleagues), 2), nsmall = 2), "±", format(round(sd(dat$imp_colleagues), 2), nsmall = 2), sep=" "))
text(bar_labels, table(dat$imp_colleagues) + 20, labels=(table(dat$imp_colleagues)))
```

##

```{r}
# Ease

bar_labels <- barplot(table(dat$easy_colleagues), ylim = c(0, max(table(dat$easy_colleagues)+30)), main = "Colleagues: Ease", xlab = paste("Mean ± stddev:", format(round(mean(dat$easy_colleagues), 2), nsmall = 2), "±", format(round(sd(dat$easy_colleagues), 2), nsmall = 2), sep=" "))
text(bar_labels, table(dat$easy_colleagues) + 20, labels=(table(dat$easy_colleagues)))
```

##

```{r}
# Knowledge

bar_labels <- barplot(table(dat$knol_colleagues), ylim = c(0, max(table(dat$knol_colleagues)+30)), main = "Colleagues: Knowledge", xlab = paste("Mean ± stddev:", format(round(mean(dat$knol_colleagues), 2), nsmall = 2), "±", format(round(sd(dat$knol_colleagues), 2), nsmall = 2), sep=" "))
text(bar_labels, table(dat$knol_colleagues) + 20, labels=(table(dat$knol_colleagues)))
```

</section>

---

<section><h1>Everything beyond this point is WIP!</h1>

## Age distributions
```{r, echo=FALSE}
age_breaks_ten <- seq(25, 85, by=10)
agegroups_count <- table(cut(dat$age, age_breaks_ten, right = F))

print(age_breaks_ten)
print(agegroups_count)

bar_labels <- barplot(agegroups_count, ylim = c(0, max(agegroups_count+30)), main = "Age distribution by decade", names.arg=c("25-34", "35-44", "45-54", "55-64", "65-74", "75-84"))
text(bar_labels, agegroups_count + 10, labels=agegroups_count)
```

##

```{r, echo=FALSE}
dat.25 <- filter(dat, age >= "25", age < "35")
dat.35 <- filter(dat, age >= "35", age < "45")
dat.45 <- filter(dat, age >= "45", age < "55")
dat.55 <- filter(dat, age >= "55", age < "65")
dat.65 <- filter(dat, age >= "65", age < "75")
dat.75 <- filter(dat, age >= "75", age < "85")
```

##

```{r, echo=FALSE}
ggplot(dat, aes(x = age)) + geom_histogram(binwidth=1)
```

##

```{r, echo=FALSE}
ggplot(dat, aes(x = age)) + geom_histogram(binwidth=5)
```
</section>